Stateful Forward Reasoning

### 2025-02-03 paper.tex
examples 1 and 2

From such lemmas, we will create what we call \textit{forward rules} that represent the theorems on which we would like to reason in a forward manner.
The premises $\alpha$, $ \text{Preorder } \alpha$ , $a$, $b$, $c$, $a \leq b$ and $b \leq c$ are separated between \textit{implied hypotheses} $\text{Preorder } \alpha$,
$a$, $b$ and $c$ and \textit{maximal hypotheses} $a \leq b$ and $b \leq c$.
Maximal hypotheses are also called \textit{input hypotheses}.

For the premises $(x_1 : T_1),(x_2 : T_2),\dots,(x_n : T_n)$, the input hypotheses consist of the hypotheses for which there are no $j > i$ such that $T_j$ depends on $x_i$.

\paragraph{Rules.}
We write $r : \Phi \to \Psi$ for a \textit{rule} $r$ with input hypotheses $\Phi$ and output hypothesis $\Psi$.
For a given rule, we assign each maximal input hypothesis a unique index between $1$ and $n$.
We call theses indices \textit{slots} of $r$ and define $\text{slots}(r)$ as the set of slots of $r$.

Given a slot $i$ and hypothesis $h : T$ with $T \equiv_{[\sigma]} T_i$ (i.e. $\sigma$ is the substitution that results from unifying $T$ and $T_i$), we say that $h$ is \textit{suitable for slot $i$} and define $\text{sub}(h) ≔ \sigma$.

A \textit{shared variable} of $r$ is an input hypothesis $x_i$ such that there are at least two input hypothesis types $T_j$, $T_k$ in which $x_i$ occurs.
A shared variable is necessarily a non-maximal input hypothesis.
We also write $?x_i$ to emphasise that we interpret $x_i$ as a variable (and, in the code, as a metavariable).
The set $s_r(i)$ of \textit{shared variables of a slot} $i$ of $r$ are those shared variables $?x$ that occur in $T_i$ and in at least one slot $j < i$.
If the rule in question is clear from the context, we write $s(i)$.

%The typical behavior is that running the procedure in a context containing suitable
%candidates for the maximal hypotheses will add the \textit{result} of the forward
%rule $a < c$ to the context.
%The result is also called the \textit{output hypothesis}.
%If we find matching hypotheses for all maximal input hypotheses in the local context,
%we can apply the forward rule, since each non-maximal input hypothesis is uniquely determined
%by maximal hypotheses.

A \textit{match} $M$ for $r$ is a partial map from slots of $r$ to hypotheses in the context.
It must satisfy the following conditions:
\begin{itemize}
  \item For each slot $i$ of $r$, if $M(i)$ is defined, then $M(i) \equiv_{[\sigma_i]} T_i$ with a substitution $\sigma_i$.
        The domain of this substitution is a subset of the non-maximal input hypotheses.
  \item For all slots $i$ and $j$ of $r$ such that $M(i)$ and $M(j)$ are defined, $\sigma_i$ and $\sigma_j$ are compatible.
  \item $M$ is downward-closed: if $M(i)$ is defined, then for any slot $j < i$ of $r$, $M(j)$ is also defined.
\end{itemize}

@@

\paragraph{Matches.}
Compatible hypotheses are hypotheses such that the instantiation of their shared variables are the same.
For example, the hypotheses $h_1 : a \leq b$ and $h_2 : b \leq c$ are compatible because they are instantiations of $x \leq y$ and $y \leq z$ where the instantiation of their shared variable $\{y\}$, is the same $\{b\}$.
A complete match is a set of hypotheses such that each of these hypothesis corresponds to exactly one slot and they are all compatible.
We proceed in the following way to find a complete match.

First, we order the slots such that each slot has at least one variable in common with the previous slots.
This can be proven to always be possible.
\xcom{We can add the proof if we want to / have space.}

@@

\paragraph{Example 1.}
\xcom{WIP}
To demonstrate the typical behavior of the algorithm we consider the following example which has \textbf{Lemma 1} as a rule and the hypotheses $(h_1 : 300 \leq 301)$ and $(h_2 : 301 \leq 302)$. \xcom{format this.}
We will consider the hypothesis $h_2$ first for demonstration purposes.

At the beginning, for the rule \textit{le\_trans}, we notice that the only shared variable is ${y}$.
Now the hypothesis $(h_2)$ unifies both with slot 1 $x \leq y$ and slot 2 $y \leq z$.
This updates the map $\eta_y(y \mapsto 302, 1)$ with $\{h_2\}$ and $\eta_y(y \mapsto 301, 2)$ with $\{h_2\}$.

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
  \node[xshift = 0] {$\eta_y$}
    child[xshift = -20] {node {1}
      child {node {$302$}
        child {node {$\{h_2\}$}}}
      child {node {$301$}
        child {node {$\{h_1\}$}}}}
    child[xshift = 20] {node {2}
      child {node {$301$}
        child {node {$\{h_2\}$}}}
      child {node {$300$}
        child {node {$\{h_1\}$}}}};
  \node[xshift = 90] {$\eta_y$}
    child[xshift = -20] {node {1}
      child {node {$302$}
        child {node {$\{h_2\}$}}}
      child {node {$301$}
        child {node {$\{h_1\}$}}}}
    child[xshift = 20] {node {2}
      child {node {$301$}
        child {node {$\{h_2\}$}}}
      child {node {$300$}
        child {node {$\{h_1\}$}}}};
  \node[xshift=200] {$\mu_y$}
    child {node[xshift=-20] {1}
      child {node {$302$}
        child {node {$\{\{1 \mapsto h_2\}\}$}}}
      child {node {$301$}
        child {node {$\{\{1 \mapsto h_1\}\}$}}}}
    child {node[xshift=20] {2}
      child {node {$301$}
        child {node[yshift=-2em] {$\{\{1 \mapsto h_1, 2 \mapsto h_2\}\}$}}}};
\end{tikzpicture}

\paragraph{Example 2.}
We now consider another example with more than one shared variable.
Let $T_1, T_2$ be two proposition that both depend on the variable $x$ and $y$.
For simplicity, the rule considered will be $T_1(x,y) \to T_2(x,y) \to R$.
The hypotheses added to the context will be, in order, $h_1 : T_2(a,a)$, $h_2 : T_1(a,b)$ and $h_3 : T_2(a,b)$.

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
  \node[xshift = 0] {$\eta_x$}
    child[xshift = -20] {node {1}
      child {node {$a$}
        child {node {$\{ \}$}}}
      child {node {$b$}
        child {node {$\{ \}$}}}}
    child[xshift = 20] {node {2}
      child {node {$a$}
        child {node {$\{h_1\}$}}}
      child {node {$b$}
        child {node {$\{ \}$}}}};
  \node[xshift = 90] {$\eta_y$}
    child[xshift = -20] {node {1}
        child {node {$a$}
            child {node {$\{ \}$}}}
        child {node {$b$}
            child {node {$\{ \}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        child {node {$b$}
            child {node {$\{ \}$}}}};
\end{tikzpicture}

At this point, there are no matches so there is nothing to extend.
We add the next hypothesis in the queue to the variable maps.

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
\node[xshift = 0] {$\eta_x$}
    child[xshift = -20] {node {1}
        child {node {$a$}
        child {node {$\{h_2\}$}}}
        child {node {$b$}
        child {node {$\{ \}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
        child {node {$\{h_1\}$}}}
        child {node {$b$}
        child {node {$\{ \}$}}}};
\node[xshift = 90] {$\eta_y$}
    child[xshift = -20] {node {1}
        child {node {$a$}
            child {node {$\{ \}$}}}
        child {node {$b$}
            child {node {$\{ h_2\}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        child {node {$b$}
            child {node {$\{ \}$}}}};
\end{tikzpicture}

As this hypothesis is in the first slot, we create the corresponding level 1 partial match $m_0 = {(h_2 : T_1(a, b))}$ for which we check for suitable hypotheses that could extend the match.
Note that  $s(lvl(m_0) + 1) = s(2)$ which is the intersection of the variables appearing in the slots $i < 2$ with the variables appearing in slot $2$.
Thus $s(2) = \{a,b\}$ and we compute:
\[
  \bigcap_{x\in s(lvl(m_0) + 1)} \eta_x \left(i + 1, \sigma_{m_0}(x)\right) = \eta_x(2, a) \cap \eta_y(2, b) = \emptyset
\]
Since the set is empty, the match can not be currently extended.
We turn to the remaining hypothesis in the queue and add it to the maps.

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
\node[xshift = 0] {$\eta_x$}
    child[xshift = -20] {node {1}
        child {node {$a$}
        child {node {$\{h_2\}$}}}
        child {node {$b$}
        child {node {$\{ \}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
        child {node {$\{h_1, h_3\}$}}}
        child {node {$b$}
        child {node {$\{ \}$}}}};
\node[xshift = 90] {$\eta_y$}
    child[xshift = -20] {node {1}
        child {node {$a$}
            child {node {$\{ \}$}}}
        child {node {$b$}
            child {node {$\{h_2\}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        child {node {$b$}
            child {node {$\{h_3\}$}}}};
\end{tikzpicture}

We are now ready to look for potential matches that could be extended by $h_3$.
In this particuliar case, since $h_3$ is in slot $2$ we look for $s(2) = \{a,b\}$.
Then,
\[
\bigcap_{x\in s(2)} \mu_x \left(i - 1, \sigma_h(x)\right) = \mu_x(1, a) \cap \mu_y(1, b) = \{m_0\}
\]

@@

\xcom{Here are some notes about the implementation.}
The index used to match hypotheses with their matching rules are implemented as discrimination trees.
A discrimination tree is a trie-like data structure that maps expressions to arbitrary data.
In~\cite{TODO}, it is shown to be good solution for the problem of finding terms (here the types of hypotheses) that might unify with other terms (here the types of the input hypotheses).
In Lean 4, discrimination trees are also used to index typeclass instances and simplifier lemmas.

@@

One of the more costly operation that we need to run multiple times is the unification \xcom{defeq?} between the hypotheses and the types of the input hypotheses associated to rules.
We implement a variant of Lean's unification process as follows: the idea is that we wish to normalize the types as much as possible.

>>>>

From such lemmas, we will create what we call \textit{forward rules} that represent the theorems on which we would like to reason in a forward manner.
The premises $\alpha$, $ \text{Preorder } \alpha$ , $a$, $b$, $c$, $a \leq b$ and $b \leq c$ are separated between \textit{implied premises} $\text{Preorder } \alpha$,
$a$, $b$ and $c$ and \textit{maximal premises} $a \leq b$ and $b \leq c$.
Maximal premises are also called \textit{input premises}.

For the premises $(x_1 : T_1),(x_2 : T_2),\dots,(x_n : T_n)$, the input premises consist of the premises for which there are no $j > i$ such that $T_j$ depends on $x_i$.

\paragraph{Rules.}
We write $r : \Phi \to \Psi$ for a \textit{rule} $r$ with input premises $\Phi$ and output $\Psi$.
For a given rule, we assign each maximal input premise a unique index between $1$ and $n$.
We call theses indices \textit{slots} of $r$ and define $\text{slots}(r)$ as the set of slots of $r$.

Given a slot $i$ and hypothesis $h : T$ in the context $\Gamma$ with $T \equiv_{[\sigma]} T_i$ (i.e. $\sigma$ is the substitution that results from unifying $T$ and $T_i$), we say that $h$ is \textit{suitable for slot $i$} and define $\text{sub}(h) ≔ \sigma$.

A \textit{shared variable} of $r$ is an input premise $x_i$ such that there are at least two input premise types $T_j$, $T_k$ in which $x_i$ occurs.
A shared variable is necessarily a non-maximal input premise.
We also write $?x_i$ to emphasise that we interpret $x_i$ as a variable (and, in the code, as a metavariable).
The set $s_r(i)$ of \textit{shared variables of a slot} $i$ of $r$ are those shared variables $?x$ that occur in $T_i$ and in at least one slot $j < i$.
If the rule in question is clear from the context, we write $s(i)$.

%The typical behavior is that running the procedure in a context containing suitable
%candidates for the maximal hypotheses will add the \textit{result} of the forward
%rule $a < c$ to the context.
%The result is also called the \textit{output hypothesis}.
%If we find matching hypotheses for all maximal input premises in the local context,
%we can apply the forward rule, since each non-maximal input premise is uniquely determined
%by maximal hypotheses.

A \textit{match} $M$ for $r$ is a partial map from slots of $r$ to hypotheses in the context.
It must satisfy the following conditions:
\begin{itemize}
  \item For each slot $i$ of $r$, if $M(i)$ is defined, then $M(i) \equiv_{[\sigma_i]} T_i$ with a substitution $\sigma_i$.
        The domain of this substitution is a subset of the non-maximal input premises.
  \item For all slots $i$ and $j$ of $r$ such that $M(i)$ and $M(j)$ are defined, $\sigma_i$ and $\sigma_j$ are compatible.
  \item $M$ is downward-closed: if $M(i)$ is defined, then for any slot $j < i$ of $r$, $M(j)$ is also defined.
\end{itemize}

@@

\paragraph{Matches.}
Compatible hypotheses are hypotheses such that the instantiation of their shared variables are the same.
For example, the hypotheses $h_1 : a \leq b$ and $h_2 : b \leq c$ are compatible because they are instantiations of $x \leq y$ and $y \leq z$ where the instantiation of their shared variable $\{y\}$, is the same $\{b\}$.
A complete match is a set of hypotheses such that each of these hypothesis corresponds to exactly one slot and they are all compatible.

Given a rule, we will partition its premises into subsets for which premises share a dependency.
To be precise, consider the graph for which the input premises $T_i$ of $r$ are the vertices and whose edges are precisely the pairs of vertices $(T_i, T_j)$ for which $v(T_i) \cap v(T_j) \neq \emptyset$.
The subsets of premises of $r$ that share a dependency are precisely the connected components of this graph.
We call these subsets of premises \textit{cluters}.

%I think, if we have space, we could also add the following characterization :
%Another characterization is the partition with the property that for two parts $C_1$ and $C_2$, $\bigcup_{T_i \in C_1} v(T_i) \cap \bigcup_{T_i \in C_2} v(T_i) = \emptyset$, and for which there do not exists a partition with the same property that has more parts.

One particularity of clusters is that, for two clusters $C_1$ and $C_2$,
$\bigcup_{T_i \in C_1} v(T_i) \cap \bigcup_{T_i \in C_2} v(T_i) = \emptyset$.
This means that compatibility need only be considered within clusters as there are no shared variables between distinct clusters.
In other words, premises in a cluster are independent of the premises in the other clusters.
Consequently, in what follows, we will assume without loss of generality that there is only one cluster.
The process we follow would simply be repeated on each cluster.

We proceed in the following way to find a complete match:
First, we order the slots such that each slot has at least one variable in common with the previous slots.
This can be proven to always be possible.
\xcom{We can add the proof if we want to / have space.}

@@

\paragraph{Example 1.}
\xcom{This is quite verbose, maybe I should use a more systematic approach.}
To demonstrate the typical behavior of the algorithm we consider the following example which has \textbf{Lemma 1} as a rule and for which we add, in order, the hypotheses $h_1 : 301 \leq 302$ and $h_2 : 300 \leq 301$.
Note that this order is choosed only for demonstration purposes.

At the beginning, for the rule \textit{le\_trans}, we notice that the only shared variable is ${y}$.
Now the hypothesis $h_1 : 301 \leq 302$ unifies both with slot 1: $x \leq y$ and slot 2: $y \leq z$.
This updates the map $\eta_y(1, y \mapsto 302)$ with $\{h_1\}$.
An update in slot 1 also marks the begining of a new partial match and we update de match map $\mu_y$ acordingly.
Note that after the creation of a match of level 1, we look at the hypotheses in slot 2 and check if they can extend the match to level 2. However, the instantiation of the current hypotheses in slot 2 are inconsistent with $y \mapsto 302$, so nothing can be extended.
Finally, as the hypothesis also matches slot 2, we update the map $\eta_y(2, y \mapsto 301)$ with $\{h_1\}$.
After these updates, the maps have the following state :

\begin{center}
\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
  \node[xshift = 0] {$\eta_y$}
    child[xshift = -20] {node {1}
      child {node {$302$}
        child {node {$\{h_1\}$}}}
    %  child {node {$301$}
    %    child {node {$\{\}$}}}
        }
    child[xshift = 20] {node {2}
      child {node {$301$}
        child {node {$\{h_1\}$}}}
    %  child {node {$300$}
    %    child {node {$\{\}$}}}
    };
  \node[xshift=160] {$\mu_y$}
    child {node[xshift=-20] {1}
      child {node {$302$}
        child {node {$\{[h_1]\}$}}}
    %  child {node {$301$}
    %    child {node {$\{\}$}}}
    }
    %child {node[xshift=20] {2}
    %  child {node {$301$}
    %    child {node[yshift=-2em] {$\{\}$}}}};
\end{tikzpicture}
\end{center}

The next step is to add $h_2 : 300 \leq 301$ to the maps.
Again, it unifies with both slots 1 and 2; we start by updating the map $\eta_y(1, y \mapsto 301)$ with $\{h_2\}$.
Again, an update in slot 1 causes the creation of a level 1 match.
We then check whether there are suitable hypotheses in slot 2 with
with a compatible instantiation of $y$; indeed we find $\eta_y(2, y \mapsto 301) = \{h_1\}$.
This update the match map with the level 2 match $[h_2,h_1]$.
This is a complete match.
The new hypothesis $h_3 : 300 \leq 302$ is added to the context.

The hypothesis $h_2$, also matches with slot 2.
This causes the update $\eta_y(2, y \mapsto 300)$ with $\{h_2\}$. 
At this point, we need to check for level 1 that would be compatible, there are none and the work with this hypothesis is done.

\begin{center}
    \begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
      \node[xshift = 0] {$\eta_y$}
        child[xshift = -20] {node {1}
          child {node {$302$}
            child {node {$\{h_1\}$}}}
          child {node {$301$}
            child {node {$\{h_2\}$}}}
            }
        child[xshift = 20] {node {2}
          child {node {$301$}
            child {node {$\{h_1\}$}}}
          child {node {$300$}
            child {node {$\{h_2\}$}}}
        };
      \node[xshift=160] {$\mu_y$}
        child {node[xshift=-20] {1}
          child[xshift=-10] {node {$302$}
            child {node {$\{[h_1]\}$}}}
          child[xshift=10] {node {$301$}
            child {node {$\{[h_2]\}$}}}
        }
        child {node[xshift=20] {2}
          child {node {$301$}
            child {node {$\{[h_2,h_1]\}$}}}};
    \end{tikzpicture}
    \end{center}

\paragraph{Example 2.}
We now consider another example with more than one shared variable.
Let $T_1, T_2$ be two proposition that both depend on the variable $x$ and $y$.
For simplicity, the rule considered will be $T_1(x,y) \to T_2(x,y) \to R$.
The hypotheses added to the context will be, in order, $h_1 : T_2(a,a)$, $h_2 : T_1(a,b)$ and $h_3 : T_2(a,b)$.
First, $h_1 : T_2(a,a)$ is considered :

% NOTE : remove comments to get full trees.
\begin{tikzpicture}[level distance=12mm, scale=0.50]
    \node{}
    \node[xshift = 30] {$\eta_x$}
        %child[xshift = -20] {node {1}
        %  child {node {$a$}
        %    child {node {$\{ \}$}}}
        %  child {node {$b$}
        %    child {node {$\{ \}$}}}}
        child[xshift = 40] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        % child {node {$b$}
        %    child {node {$\{ \}$}}}
        };
    \node[xshift = 110] {$\eta_y$}
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        child[xshift = 40] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        };
    \node[xshift = 190, yshift = -25] {$\mu_x$}
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 270, yshift = -25] {$\mu_y$}
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
\end{tikzpicture}

At this point, there are no matches so there is nothing to extend.
We add the next hypothesis in the queue, that is $h_2 : T_1(a,b)$, to the variable maps :

% NOTE : uncomment to get full trees.
\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
    \node[xshift = 0] {$\eta_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{h_2\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
            }
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
            };
    \node[xshift = 80] {$\eta_y$}
        child[xshift = -20] {node {1}
            %child {node {$a$}
            %    child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{ h_2\}$}}}}
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
                };
    \node[xshift = 160, yshift = 0] {$\mu_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{[h_2]\}$}}}
            %child {node {$b$}
            %   child {node {$\{ \}$}}}
                }
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 240, yshift = 0] {$\mu_y$}
        child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{[h_2]\}$}}}
        }
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
\end{tikzpicture}

As this hypothesis is in the first slot, we create the corresponding level 1 partial match $m_0 = {(h_2 : T_1(x, y))}$ for which we check for suitable hypotheses that could extend the match.
Note that  $s(lvl(m_0) + 1) = s(2)$ which is the intersection of the variables appearing in the slots $i < 2$ with the variables appearing in slot $2$.
Thus $s(2) = \{x,y\}$ and we compute:
\[
  \bigcap_{x\in s(lvl(m_0) + 1)} \eta_x \left(i + 1, \sigma_{m_0}(x)\right) = \eta_x(2, a) \cap \eta_y(2, b) = \emptyset
\]
Since the set is empty, the match can not be currently extended.
We turn to the remaining hypothesis $h_3 : T_2(a,b)$ in the queue and add it to the maps:

\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
\node[xshift = 0] {$\eta_x$}
    child[xshift = -20] {node {1}
        child {node {$a$}
            child {node {$\{h_2\}$}}}
        %child {node {$b$}
        %    child {node {$\{ \}$}}}
        }
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1, h_3\}$}}}
        %child {node {$b$}
        %    child {node {$\{ \}$}}}
        };
\node[xshift = 80] {$\eta_y$}
    child[xshift = -20] {node {1}
        %child {node {$a$}
        %    child {node {$\{ \}$}}}
        child {node {$b$}
            child {node {$\{h_2\}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        child {node {$b$}
            child {node {$\{h_3\}$}}}};
    \node[xshift = 160, yshift = 0] {$\mu_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{[h_2]\}$}}}
            %child {node {$b$}
            %   child {node {$\{ \}$}}}
                }
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{[h_2,h_3]\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        };
    \node[xshift = 240, yshift = 0] {$\mu_y$}
        child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{[h_2]\}$}}}
        }
        child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
            child {node {$b$}
                child {node {$\{[h_2,h_3]\}$}}}
        };
\end{tikzpicture}

We are now ready to look for potential matches that could be extended by $h_3$.
In this particuliar case, since $h_3$ is in slot $2$ we look for $s(2) = \{x,y\}$.
Then,
\[
\bigcap_{x\in s(2)} \mu_x \left(i - 1, \sigma_h(x)\right) = \mu_x(1, a) \cap \mu_y(1, b) = \{m_0\}
\]

@@

\xcom{Here are some notes about the implementation.}
The index used to match hypotheses with their matching rules are implemented as discrimination trees.
A discrimination tree is a trie-like data structure that maps expressions to arbitrary data.
In~\cite{TODO}, it is shown to be good solution for the problem of finding terms (here the types of hypotheses) that might unify with other terms (here the types of the input premises).
In Lean 4, discrimination trees are also used to index typeclass instances and simplifier lemmas.

@@

One of the more costly operation that we need to run multiple times is the unification \xcom{defeq?} between the hypotheses and the types of the input premises associated to rules.
We implement a variant of Lean's unification process as follows: the idea is that we wish to normalize the types as much as possible.
